---
title: "Data Cleaning Steps"
author: "Syafeeq Alias"
date: "12/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will read the CSV file we get from MOF website and transform & clean it to tidy data.

First we will clean up the Federal Government Revenue Data

```{r}
library(tidyr)

#Federal Government Revenue Data
FedGovRev19702017DF <- read.csv("Fed Gov Revenue 1970-2017.csv")

#check the data file
names(FedGovRev19702017DF)
dim(FedGovRev19702017DF)
head(FedGovRev19702017DF)

#initial data frame was Entry, Year 1, Year 2, ..., Year N
#tidy up the data file so it goes to the right structure, i.e. Entry, Year, Value
Tidy_FedGovRev19702017DF <- gather(FedGovRev19702017DF, "Year", "Value", 3:49)
head(Tidy_FedGovRev19702017DF)
```

The data cleaning steps consists of the following steps

1. Remove X from the Year column
2. Substitute '-' with 0
3. Change the Year column to factor
4. Change the Value column to numeric
5. Change NA Value to 0

Once cleaned, we save the data frame to anoter CSV file for use

```{r}
Tidy_FedGovRev19702017DF$Year <- gsub('X','',Tidy_FedGovRev19702017DF$Year)
Tidy_FedGovRev19702017DF$Value <- gsub('-','0',Tidy_FedGovRev19702017DF$Value)
Tidy_FedGovRev19702017DF <- transform(Tidy_FedGovRev19702017DF, Year = as.factor(Year))
Tidy_FedGovRev19702017DF <- transform(Tidy_FedGovRev19702017DF, Value = as.numeric(Value))
Tidy_FedGovRev19702017DF$Value[is.na(Tidy_FedGovRev19702017DF$Value)] <- 0

#check data is now in right format
head(Tidy_FedGovRev19702017DF)
sapply(Tidy_FedGovRev19702017DF, class)

#write the tidy data to a csv file
write.csv(Tidy_FedGovRev19702017DF, "Tidy_FedGovRev.csv", row.names = FALSE)
```

We will do the same data cleaning steps for the data file for Federal Government Development Expenditure

```{r}
#Federal Government Development Expenditure Data
FedGovDevex19702017DF <- read.csv("Fed Gov Devex 1970-2017.csv")
head(FedGovDevex19702017DF)

Tidy_FedGovDevex <- gather(FedGovDevex19702017DF, "Year", "Value", 3:50)
head(Tidy_FedGovDevex)

Tidy_FedGovDevex$Year <- gsub('X','',Tidy_FedGovDevex$Year)
Tidy_FedGovDevex$Value <- gsub('-','0',Tidy_FedGovDevex$Value)
Tidy_FedGovDevex <- transform(Tidy_FedGovDevex, Year = as.factor(Year))
Tidy_FedGovDevex <- transform(Tidy_FedGovDevex, Value = as.numeric(Value))
Tidy_FedGovDevex$Value[is.na(Tidy_FedGovDevex$Value)] <- 0

#check data is now in right format
head(Tidy_FedGovDevex)
sapply(Tidy_FedGovDevex, class)

#write the tidy data to a csv file
write.csv(Tidy_FedGovDevex, "Tidy_FedGovDevex.csv", row.names = FALSE)

```

Repeat data cleaning steps for Federal Government Opex

```{r}
#Federal Government Opeerating Expenditure Data
FedGovOpex19702017DF <- read.csv("Fed Gov Opex 1970-2017.csv")
head(FedGovOpex19702017DF)

Tidy_FedGovOpex <- gather(FedGovOpex19702017DF, "Year", "Value", 3:50)
head(Tidy_FedGovOpex)

Tidy_FedGovOpex$Year <- gsub('X','',Tidy_FedGovOpex$Year)
Tidy_FedGovOpex$Value <- gsub('-','0',Tidy_FedGovOpex$Value)
Tidy_FedGovOpex <- transform(Tidy_FedGovOpex, Year = as.factor(Year))
Tidy_FedGovOpex <- transform(Tidy_FedGovOpex, Value = as.numeric(Value))
Tidy_FedGovOpex$Value[is.na(Tidy_FedGovOpex$Value)] <- 0

#check data is now in right format
head(Tidy_FedGovOpex)
sapply(Tidy_FedGovOpex, class)

#write the tidy data to a csv file
write.csv(Tidy_FedGovOpex, "Tidy_FedGovOpex.csv", row.names = FALSE)

```